---
title: "Municipio Spatial"
author: "José A. Bartolomei-Díaz, PhD - Outcome Project"
date: "June 5, 2015"
output: pdf_document
---

```{r rm3}
rm(list = ls(all = TRUE))
```


library

```{r library3}
library(ggplot2)
library(ggmap)
library(sp)
library(rgeos)
library(maptools)
library(xtable)
```


Load managed Spatial Polygon of Puerto Rico
```{r load3}

load("/media/truecrypt2/ORP2/Spatial/R_Geo_Layers/Puerto_Rico/PR_Muni_Map.RData")

map <- PR_Muni_Map
rm(PR_Muni_Map)
```


```{r Cases_df}
load("/media/truecrypt2/ORP2/Alzheimer_Registry/Data/Managed_Data/AlDf.RData")
```

# There are 27 records without municipalities.


Select complete cases
```{r completeCases}
### Create a data.frame excluding those rows wich AgeCut == "NA"
#AlDfB <- 	AlDf[!(is.na(AlDf$Edad)),] Extrae los NA pero solo de una columna a la vez
AlDf <- AlDf[complete.cases(AlDf[, c("Edad", "Genero")]),] 
```
Visualy check the remove Na's in Edad
```{r check3}
head(data.frame(AlDf$Edad, AlDf$AgeCut))
tail(data.frame(AlDf$Edad, AlDf$AgeCut))
table(AlDf$AgeCut)
```


Check if Gender have Na'
```{r Gender_NA}
#
table(AlDf$Genero)
#Remove those rows with gender == "" (Blanck Space)
AlDf <- AlDf[AlDf$Genero != "",]
AlDf$Genero <- factor(AlDf$Genero,
											labels = c("Female", "Male"))
table(AlDf$Genero)
```
```{r Muni_NA}
##################################################
# Check if Municipality have Na's
length(AlDf$Municipio.Recidencia)
#Remove those rows with gender == "" (Blanck Space)
AlDf <- AlDf[AlDf$Municipio.Recidencia != "",]
length(AlDf$Municipio.Recidencia)
```

```{r length}
length(table(as.factor(AlDf$Municipio.Recidencia), exclude = ""))
# Slecting zips using the table found in the internet
# AlDf <-
# 	dim(AlDf[!AlDf$Municipio.Recidencia %in% "",])


A_cases <- data.frame(table(as.factor(AlDf$Municipio.Recidencia), exclude = ""))
colnames(A_cases)	<- c("Muni", "Cases")
# There is a need to transform to character and order the muni names

A_cases
```
# Hospital Location Coverage

Load Hospital Layer
```{r Load_Hops_Layer}
HospL <- readShapeSpatial("/home/surfprjab/Dropbox/OutcomeProjectTeam/shapefile/Hospital_Locations/HOSP_2014.shp")
HospL2 <- as.data.frame(HospL)
```

Create a vector of aggregated values by municipalities
```{r CutValues}
# Create aggregated values of Cases

###########################################
### Cut 1 ####
###########################################
A_cases$Cut <- cut(A_cases$Cases,
									 breaks = c("0", "1", "100", "200",
									 					 "300", "400", Inf),
									 labels = c("0-1", "2-100", "101-200", "201-300",
									 					 "301-400", "401-1677")
									 )

summary(A_cases$Cut)

###########################################
### Cut IQR ####
###########################################
options(scipen=999)
A_cases$CutIQR <- cut(A_cases$Cases, breaks = c(0, quantile(A_cases$Cases)),
									labels = c("0-2", "2-47.2", "47.3-111", "112-202",
										 paste("203-", max(A_cases$Cases), sep = ""))
											)

summary(A_cases$CutIQR)
```

A row unique ID will be needed to merge both data.frame. This is becuase there are some differences in municipalities names in the data frames.
```{r IDnum}
head(A_cases)
A_cases$IDnum <- seq(1:78)
```

Verify colums to merge
```{r Verify2Merge}
data.frame(map[!duplicated(map$id),][, 1:2], A_cases[, c("Muni", "IDnum")])
```


An id per municipality because id name did not matched between the Spatial Data and the AlDf. 
```{r un_Muni}
# unique Municiaplities
un_Muni <- data.frame(MuniID = unique(map$id))
un_Muni$IDnum <- seq(1:78)
# 
# map <- merge(x = map, y = un_Muni, by.x = "IDnum",
# 												 by.y = "IDnum")

# verificar que los municipios tengan el mismo id para poder merge.
data.frame(map[!duplicated(map$id),][, 1:2], un_Muni[, c("MuniID", "IDnum")])

map <- merge(x = map, y = A_cases[, -1], by.x = "IDnum",
												 by.y = "IDnum")
head(map)

```

```{r MapMuni_Cases_Cut6, fig.width=15, fig.height=10, fig.path="/media/truecrypt2/ORP2/Alzheimer_Registry/Figures/"}

MapMuni_Cases_Cut6 <- 
	ggplot(data = map) +
		geom_polygon(aes(x = long, y = lat, group = id,
										 fill = Cut)) + 
#geom_text(data = MuniID, aes(x = long, y = lat, label = id)) +
	
			geom_polygon(aes(x = long, y = lat, group = group, label = "id"),
								 data = map, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	coord_map() + 
	
#	ggtitle("Alzheimer cases by municipality, 2008-2013") +
	# Include total number of caes
	geom_text(data = NULL, x = -65.50, y = 18.46, size = 6,
						label = paste("N = ", format(nrow(AlDf),
																				 big.mark = ","),
													sep = "")) +
	# Include mean and SDeviation
	geom_text(data = NULL, x = -65.57, y = 17.96, size = 4, 
						label = paste("mean = ",  round(mean(A_cases$Cases), 2),
													" ; ",
													"sd = ",  round(sqrt(var(A_cases$Cases)), 2),
																					 sep = "")) +
	geom_point(data = HospL2,
											aes(x = LONGITUDE, y = LATITUD),
											color = "Yellow") +
	
	scale_fill_brewer(palette = 5, 
										guide = guide_legend(title = "Number of Cases")) +
	labs(x = "Longitude", y = "Latitude") 

MapMuni_Cases_Cut6

```

### Quartile Muni Map

```{r MapMuni_Cases_CutIQR, fig.width=15, fig.height=10, fig.path="/media/truecrypt2/ORP2/Alzheimer_Registry/Figures/"}

MapMuni_Cases_CutIQR <- 
	ggplot(data = map) +
		geom_polygon(aes(x = long, y = lat, group = id,
										 fill = CutIQR)) + 
#geom_text(data = MuniID, aes(x = long, y = lat, label = id)) +
	
			geom_polygon(aes(x = long, y = lat, group = group, label = "id"),
								 data = map, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	coord_map() + 
	
	ggtitle("Alzheimer cases by municipality, 2008-2013") +
	geom_text(data = NULL, x = -65.35, y = 18.46, 
						label = paste("N = ", format(nrow(AlDf),
																				 big.mark = ",")
						)
	) +
	geom_point(data = HospL2,
											aes(x = LONGITUDE, y = LATITUD),
											color = "Yellow") +
	
	scale_fill_brewer(palette = 5) +
	labs(x = "Longitude", y = "Latitude")

MapMuni_Cases_CutIQR
```

## Map With Info of Cases by Gender

```{r Agen_Cases}
Agen_cases <- data.frame(table(as.factor(AlDf$Municipio.Recidencia),
															 AlDf$Genero, exclude = ""))

colnames(Agen_cases) <- c("Muni", "Gender", "Cases")

Agen_cases
#################################################################
# Because municipalities ids do not match run the following

Agen_cases$IDnum <- rep(seq(1:78), 2)

SpatialDataMuniC <- 
	map[, !(names(map) %in% c("Cases", "Cut", "Cut2"))]
	
# verificar que los municipios tengan el mismo id para poder merge.
data.frame(SpatialDataMuniC[!duplicated(SpatialDataMuniC$id),][, 1:2], Agen_cases[, c("Muni", "IDnum")])

####
SpatialDataMuniC <- merge(x = SpatialDataMuniC, y = Agen_cases,
													by.x = "IDnum", by.y = "IDnum")
head(SpatialDataMuniC)

# Create aggregated values of Cases
SpatialDataMuniC$Cut <- cut(SpatialDataMuniC$Cases,
									 breaks = c("0", "1", "100", "200",
									 					 "300", "400", Inf),
									 labels = c("0-1", "2-100", "101-200", "201-300",
									 					 "301-400", "401+")
									 )
# 
# SpatialDataMuniC$Cut2 <- cut(SpatialDataMuniC$Cases,
# 														 breaks = c("0", "1", "100", "300", "500", Inf),
# 														 labels = c("0-1", "2-100", "101-300", "301-500",
# 														 					 "501+"))

```

## Map A cases facet by gender

```{r MapMuni_CasesGender, fig.width=15, fig.height=10, fig.path="/media/truecrypt2/ORP2/Alzheimer_Registry/Figures/"}

MapMuni_CasesGender <- 
	ggplot(data = SpatialDataMuniC) +
		geom_polygon(aes(x = long, y = lat, group = group,
										 fill = Cut)) + 
#geom_text(data = MuniID, aes(x = long, y = lat, label = id)) +
	
			geom_polygon(aes(x = long, y = lat, group = group),
								 data = SpatialDataMuniC, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	coord_map() + 
	facet_wrap(~ Gender, ncol = 1) +

#	ggtitle("Alzheimer cases by municipality and gender, 2008-2013") + 
	geom_text(aes(x, y, label=lab),
						data = data.frame(x = -65.35, y = 18.46, 
														lab = c(paste("N = ", 
																	format(nrow(subset(AlDf,
																										 Genero == "Femenino")),
 																				 big.mark = ",")),
																	paste("N = ", 
																	format(nrow(subset(AlDf,
																										 Genero == "Masculino")),
 																				 big.mark = ","))),
														Gender = c("Female", "Male")), 
													vjust=1) +
	
	scale_fill_brewer(palette = 5,
										guide = guide_legend(title = "Number of Cases")) +
	labs(x = "Longitude", y = "Latitude")


MapMuni_CasesGender

```

```{r tableCheckCut}
table(A_cases$Cut)
```

```{r SaveObjects3}
save(MapMuni_Cases_Cut6, MapMuni_Cases_CutIQR, MapMuni_CasesGender, A_cases,
file = "/media/truecrypt2/ORP2/Alzheimer_Registry/Objects/Map_Cases_Objects.Rdata")
```

```{r source_04}
# # Convert from .Rnw to .R
# knitr::knit(input = "/media/truecrypt2/ORP2/Alzheimer_Registry/04_Observed_Cases.Rmd",
# 						output = "/media/truecrypt2/ORP2/Alzheimer_Registry/04_Observed_Cases.R",
# 						tangle = TRUE)
# 
# # Source the Coverted R
# source("/media/truecrypt2/ORP2/Alzheimer_Registry/04_Observed_Cases.R")
```
