---
title: "Standardized Incidence Ratio by Municipalities, Puerto Rico"
author: "José A. Bartolomei-Díaz, PhD - Outcome Project"
output: html_document
---

This document calculate the SIR using the data prepared in the Observed_Cases.Rmd and the Spatial_Population.Rmd.

# Remove Un-needed objects
```{r rm_Objects}
rm(list = ls(all = TRUE))
```
# Load needed objects created in previous scripts
```{r load_Obs_Cases, echo=FALSE, results='hide'}

# Observed Cases
load("./Objects/Observed_Cases.RData")
```
# Load Census Population for analysis

Load Census population arrangements from: /Population_Estimates/Census/Pop_Estimates_2010.RData.

If new arrangement is needed open and edit the following file: 
./Population_Estimates/Census/Populaton_Reshape_4_Analysis.Rmd


#### Load population arrangements and select the needed to conduct the analysis.

```{r Load_popMuniSexAge, echo=FALSE, results='hide'}
load("./Population_Estimates/Census/Pop_Estimates_2010.RData")
```

# Crude Rate by Municipality

```{r Muni_Crude_Rate, echo=FALSE, results='hide'}
# Check muni order
data.frame(Obs_Muni, Pop_Muni)
# Muni_Crude_Rate
MCR <- data.frame(IDnum = seq(1:78),
															Name = Obs_Muni[, 1],
															CrRate = (Obs_Muni[, 2] / Pop_Muni[, 2]) * 10000)

#MCR <- MCR[, -2]
MCR
```

```{r CrRateCut, echo=FALSE, results='hide'}


###########################################
### Cut IQR ####
###########################################
# Create an interqualtile SIR_G cut
# The firt quartile was excluded because the small percentages of municipalities in this range and because was less than 1 per 10000

MCR$CrIQR <- cut(MCR$CrRate, breaks = c(0, quantile(MCR$CrRate)),
									labels = c("0-2.39", "2.4-14.0", "14.1-26.7",
														 "26.8-37.8", "37.9-68.4")
)

MCR[order(MCR$CrRate),]
# MCR$CrIQR <- cut(MCR$CrRate, breaks = c(0, quantile(MCR$CrRate)[-1]),
# 									labels = c("0-9.51", "9.52-25.2", "25.3-66.7",
# 														 "66.7-880")
# )
```

## Map result
## load Puerto Rico Municipality map
```{r loadMuniMap, echo=FALSE, results='hide'}
load("./Spatial_Polygons/Puerto_Rico/PR_Muni_Map.RData")

map <- PR_Muni_Map
rm(PR_Muni_Map)

library(ggplot2)
# Verify if the loaded polygons information display the desired map
ggplot(data = map) +
	geom_polygon(aes(x = long, y = lat, fill = id, alpha = .4)) +
	coord_map()
```

# Hospital Location Coverage

Load Desire Geographical Layers for analysis
```{r Load_Hops_Layer, echo=FALSE, results='hide'}
Path <- "./Spatial/R_Geo_Layers/Puerto_Rico/"
# LL = Layer List
LL <- c("C330L", "CdtL", "HospL")

for(i in 1:length(LL)){
	LLi <- LL[i]
	file <- paste(Path, LL[i], ".RData", sep = "")
	load(file = file)
}
```

## Merge Crude Rate data with spatial data
```{r df5, echo=FALSE, results='hide'}
# verificar que los municipios tengan el mismo id para poder merge.
data.frame(map[!duplicated(map$id),][, 1:2], MCR[, c("IDnum", "Name")])

map <- merge(x = map, y = MCR, by.x = "IDnum", by.y = "IDnum" )
```
# Map Crude Rate Estimates

```{r CrudeRateMap, , echo=FALSE, results='asis', fig.width=15, fig.height=10, fig.path="/Figures/"}

# Arrange the sir information with coordinates
MuniCrRate <- aggregate(cbind(long, lat) ~ CrRate, data = map,
                    FUN=function(x)mean(range(x)))

MapMuni_Crude_Rate <-
	ggplot(data = map) +
	geom_polygon(aes(x = long, y = lat, group = id, fill = CrIQR)) +
	
	geom_polygon(aes(x = long, y = lat, group = group, label = "id"),
								 data = map, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	
		#geom_text(data = MuniCrRate, aes(x = long, y = lat, label = round(CrRate, 1))) +
	
	# HospL
	geom_point(data = HospL,
						 aes(x = LONGITUDE, y = LATITUD),
						 color = "Yellow") +
# C330L
# 	geom_point(data = C330L,
# 						 aes(x = LONG, y = LAT),
# 						 color = "Blue") +
# CDTL	
# 	geom_point(data = CdtL,
# 						 aes(x = LONG_, y = LAT),
# 						 color = "Red") +
	
	coord_map() +
	
	scale_fill_brewer(palette = 5,
										guide = guide_legend(title = "Prevalence per 10,000")) +
	
#	ggtitle("Alzheimer crude rate by municipality, 2008-2013 \n Dots: Yellow = Hospitals") +
#; Blue = 330 Centers; Red = CDT"
	geom_text(data = NULL, x = -65.35, y = 18.46, 
						label = paste("N = ", format(nrow(AlDf),
																				 big.mark = ","))) +
	labs(x = "Longitude", y = "Latitude")

MapMuni_Crude_Rate

```


# SIR
```{r lib_EpiR, echo=FALSE, results='hide'}
library(epiR)
options(scipen=999)
```
# Un-adjusted Spatial Standard morbidity Ratio

# Un-adjusted Spatial Standard morbidity Ratio

# Spatial Standard morbidity Ratio adjusted by Gender
```{r un_SIR, echo=FALSE, results='hide'}
obs <- matrix(Obs_Muni[, 2], dimnames = list(Obs_Muni[, 1], ""))
obs

pop <- matrix(Pop_Muni[, 2], dimnames = list(Pop_Muni[, 1], ""))
pop

std <- sum(obs) / sum(pop)

std
## SIR using funciton epi.indirectadj
un_sir <- epi.indirectadj(obs = obs, pop = pop, std = std, units = 1)

# Create a data frame with aditional information for other procedures
un_sir <- data.frame(IDnum = seq(1:78), Muni = rownames(un_sir$smr.strata),
										un_sir = un_sir$smr.strata$est)
head(un_sir)

# Create a manual cut
un_sir$unsirCut <- cut(un_sir$un_sir, breaks = c(0, 1, 1.5, 2, 4, Inf),
										labels = c("0-1.00", "1.01-1.49", "1.50-1.99",
															 "2.00-3.99","4+")
)


```


## Merge SIR data with spatial data
```{r map_unsir_merge, echo=FALSE, results='hide'}
map <- merge(x = map, y = un_sir, by.x = "IDnum", by.y = "IDnum" )
```

## Map gender adjusted SIR estimates

```{r MapMuni_un_SIR, echo=FALSE, results='asis', fig.width=15, fig.height=10, fig.path="/Figures/"}

# Arrange the sir information with coordinates
MuniunSir <- aggregate(cbind(long, lat) ~ un_sir + id, data = map,
										 FUN=function(x)mean(range(x)))

MapMuni_un_SIR <-
	ggplot(data = map) +
	geom_polygon(aes(x = long, y = lat, group = id, fill = unsirCut)) +
	
	geom_polygon(aes(x = long, y = lat, group = group, label = "id"),
							 data = map, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	
	#geom_text(data = MuniSir, aes(x = long, y = lat, label = round(sirG, 1))) +
	
	# HospL
	geom_point(data = HospL,
						 aes(x = LONGITUDE, y = LATITUD),
						 color = "Yellow") +
	# C330L
	# 	geom_point(data = C330L,
	# 						 aes(x = LONG, y = LAT),
	# 						 color = "Blue") +
	# CDTL	
	# 	geom_point(data = CdtL,
	# 						 aes(x = LONG_, y = LAT),
	# 						 color = "Red") +
	
	coord_map() +
	
	scale_fill_brewer(palette = 5,
										guide = guide_legend(title = "Standardized Morbidity Ratio")) +
	
	#	ggtitle("Alzheimer un-adjusted Standardized Morbidity Ratio by municipality, 2008-2013 \n Dots: Yellow = Hospitals") +
	#; Blue = 330 Centers; Red = CDT"
	
	geom_text(data = NULL, x = -65.35, y = 18.46, 
						label = paste("N = ", format(nrow(AlDf),
																				 big.mark = ","))) +
	labs(x = "Longitude", y = "Latitude")

MapMuni_un_SIR
```

# Spatial Standard morbidity Ratio adjusted by Gender
```{r SIR_Sex, echo=FALSE, results='hide'}
obs <- matrix(Obs_Muni[, 2], dimnames = list(Obs_Muni[, 1], ""))
obs

pop <- as.matrix(Pop_MG[, c("Male", "Female")])
rownames(pop) <- Obs_Muni[, 1]							
pop

std <- data.frame(Male = Obs_Gen[2, "IDrow"] / Pop_Gen[1, "EST10"],
									Female = Obs_Gen[1, "IDrow"] / Pop_Gen[2, "EST10"],
									Total = sum(Obs_Gen[, "IDrow"]) / sum(Pop_Gen[, "EST10"])
)

std <- as.matrix(std)
std
## SIR using funciton epi.indirectadj
sir_G <- epi.indirectadj(obs = obs, pop = pop, std = std, units = 1)

# Create a data frame with aditional information for other procedures
sir_G <- data.frame(IDnum = seq(1:78), Muni = rownames(sir_G$smr.strata),
										sirG = sir_G$smr.strata$est)
head(sir_G)

# Create a manual cut
sir_G$sirCut <- cut(sir_G$sirG, breaks = c(0, 1, 1.5, 2, 4, Inf),
									labels = c("0-1.00", "1.01-1.49", "1.50-1.99",
														 "2.00-3.99","4+")
)

# Create an interqualtile SIR_G cut
sir_G$sirIQR <- cut(sir_G$sirG, breaks = c(0, quantile(sir_G$sirG)),
									labels = c("0-0.0179", "1.0180-0.315", "0.316-0.833", "0.834-2.2",
														 "2.3-30")
)

summary(sir_G$sirIQR)
```


## Merge SIR data with spatial data
```{r map, echo=FALSE, results='hide'}
map <- merge(x = map, y = sir_G, by.x = "IDnum", by.y = "IDnum" )
```

## Map gender adjusted SIR estimates

```{r MapMuni_adjG_SIR, echo=FALSE, results='asis', fig.width=15, fig.height=10, fig.path="/Figures/"}

# Arrange the sir information with coordinates
MuniSir <- aggregate(cbind(long, lat) ~ sirG, data = map,
                    FUN=function(x)mean(range(x)))

MapMuni_adjG_SIR <-
	ggplot(data = map) +
	geom_polygon(aes(x = long, y = lat, group = id, fill = sirCut)) +
	
	geom_polygon(aes(x = long, y = lat, group = group, label = "id"),
								 data = map, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	
		#geom_text(data = MuniSir, aes(x = long, y = lat, label = round(sirG, 1))) +
	
	# HospL
	geom_point(data = HospL,
						 aes(x = LONGITUDE, y = LATITUD),
						 color = "Yellow") +
# C330L
# 	geom_point(data = C330L,
# 						 aes(x = LONG, y = LAT),
# 						 color = "Blue") +
# CDTL	
# 	geom_point(data = CdtL,
# 						 aes(x = LONG_, y = LAT),
# 						 color = "Red") +
	
	coord_map() +
	
	scale_fill_brewer(palette = 5,
										guide = guide_legend(title = "Standardized Morbidity Ratio")) +
	
#	ggtitle("Alzheimer sex adjusted Standardized Morbidity Ratio by municipality, 2008-2013 \n Dots: Yellow = Hospitals") +
#; Blue = 330 Centers; Red = CDT"
	
	geom_text(data = NULL, x = -65.35, y = 18.46, 
						label = paste("N = ", format(nrow(AlDf),
																				 big.mark = ","))) +
	labs(x = "Longitude", y = "Latitude")

MapMuni_adjG_SIR
```

## SIR adjusted by gender and age group

```{r SIR_SexAge, echo=FALSE, results='hide'}
# obs will not be to be constructed because the same can be use.
obs <- as.matrix(Obs_MGAc[ , -which(names(Obs_MGAc) %in% c("Municipio.Recidencia"))])
rownames(obs) <- Obs_MGAc[, "Municipio.Recidencia"]							
# Chage 0 to 0.005 to avoind Inf in the SIR results
obs[obs == 0] <- 0.005
head(obs)
# Population
pop1 <- Pop_MGAc[ , -which(names(Pop_MGAc) %in% c("Male_0.19", "Female_0.19"))]
head(pop1)

pop <- as.matrix(pop1[ , -which(names(pop1) %in% c("Muni"))])
rownames(pop) <- pop1[, "Muni"]							
pop


std <- data.frame(std = c(colSums(obs)/colSums(pop), Total = sum(obs) / sum(pop)))

std <- as.matrix(std)
std
## SIR using funciton epi.indirectadj
sir_MGA <- epi.indirectadj(obs = obs, pop = pop, std = std, units = 0)

# Create a data frame with aditional information for other procedures
sir_MGA <- data.frame(IDnum = seq(1:78), Muni = rownames(sir_MGA$smr.strata),
										sirMga = sir_MGA$smr.strata$est)
head(sir_MGA)

# Create a manual cut
sir_MGA$sirMgaCut <- cut(sir_MGA$sirMga, breaks = c(0, 1, 1.5, 2, 4, Inf),
									labels = c("0-1.00", "1.01-1.49", "1.50-1.99",
														 "2.00-3.99","4+")
)

# Create an interqualtile SIR_G cut
sir_MGA$sirMgaIQR <- cut(sir_MGA$sirMga, breaks = c(0, quantile(sir_MGA$sirMga)),
									labels = c("0-0.0179", "1.0180-0.315", "0.316-0.833", "0.834-2.2",
														 "2.3-30")
)

summary(sir_MGA$sirMgaIQR)
```

## Map gender adjusted SIR estimates

### Merge SIR data with spatial data
```{r sir_MGA, echo=FALSE, results='hide'}
sir_MGA <- sir_MGA[, -2] # remove municipality columns to not repeate the column in map data when merged
map <- merge(x = map, y = sir_MGA, by.x = "IDnum", by.y = "IDnum" )
```

```{r MapMuni_adjGA_SIR, echo=FALSE, results='asis', fig.width=15, fig.height=10, fig.path="/Figures/"}

MuniSir2 <- aggregate(cbind(long, lat) ~ sirMga + id +IDnum, data = map,
                    FUN=function(x)mean(range(x)))

MapMuni_adjGA_SIR <- 
	ggplot(data = map) +
	geom_polygon(aes(x = long, y = lat, group = id, fill = sirMgaCut)) +
	
	geom_polygon(aes(x = long, y = lat, group = group, label = "id"),
								 data = map, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	
		#geom_text(data = MuniSir2, aes(x = long, y = lat, label = round(sirMga, 1))) +
	
		# HospL
	geom_point(data = HospL,
						 aes(x = LONGITUDE, y = LATITUD),
						 color = "Yellow") +
# C330L
# 	geom_point(data = C330L,
# 						 aes(x = LONG, y = LAT),
# 						 color = "Blue") +
# CDTL	
# 	geom_point(data = CdtL,
# 						 aes(x = LONG_, y = LAT),
# 						 color = "Red") +
	
	coord_map() +
	
		scale_fill_brewer(palette = 5,
										guide = guide_legend(title = "Standardized Morbidity Ratio")) +
	
#	ggtitle("Alzheimer age-sex adjusted Standardized Morbidity Ratio by municipality, 2008-2013 \n Dots: Yellow = Hospitals") +
#; Blue = 330 Centers; Red = CDT"
	geom_text(data = NULL, x = -65.38, y = 18.46, 
						label = paste("N = ", format(nrow(AlDf),
																				 big.mark = ",")
						)) +
		geom_text(data = NULL, x = -65.50, y = 18.00, size = 4,
						label = paste("mean = ", round(mean(MuniSir2$sirMga), 2), " ; ",
													"sd = ", round(sqrt(var(MuniSir2$sirMga)), 2),
													sep = "")) +
	labs(x = "Longitude", y = "Latitude")

MapMuni_adjGA_SIR
```

# Confidence Intervals by Municipality

```{r CIbyMuni}
## SIR using funciton epi.indirectadj
sir_MGA <- epi.indirectadj(obs = obs, pop = pop, std = std, units = 0)$smr.strata
sir_MGA <- data.frame(Muni = row.names(sir_MGA), sir_MGA)
```

# Significance Tests
```{r epitools}
library(epitools)

```


# Empirical Bayes Smoothing
```{r DCluster}
library(DCluster)
# Poisson-Gamma Empirical Bayes Smoothing
eb <- with(data = sir_MGA, empbaysmooth(Observed = obs, Expected = exp))
mean(eb$smthrr)
sqrt(var(eb$smthrr))

# Log Normal Smoothing
ebln <- with(data = sir_MGA, lognormalEB(Observed = obs, Expected = exp))
mean(ebln$smthrr)
sqrt(var(ebln$smthrr))

# EB Marshal
library(spdep)
ebM <- with(data = sir_MGA, EBest(n = obs, x = exp))
mean(ebM$estmm)
sqrt(var(ebM$estmm))

# Spatially Structured Model
# Load Spatial Structure
# load("/media/truecrypt2/ORP2/Spatial/PR_Spatial_Structure.RData")
# with(data = sir_MGA, EBlocal(ri = obs, ni = exp, nb = nb,
# 														 zero.policy = FALSE))

```


## Add a binary significat collumn based on CI's
```{r CreateSigVect}
sir_MGA$sig[sir_MGA$est > 1 & sir_MGA$lower > 1] <- "Significantly Higer"

sir_MGA$sig[sir_MGA$est < 1 & sir_MGA$upper < 1] <- "Significantly lower"

sir_MGA$sig[sir_MGA$est > 1 & sir_MGA$lower <= 1 |
						sir_MGA$est < 1 & sir_MGA$upper >= 1] <- "Not Significant"

table(sir_MGA$sig)
head(sir_MGA)	
```

```{r sir_MGA_df}
sir_MGA <- data.frame(IDnum =  seq(1:78), sir_MGA)
```


```{r Graph_CIbyMuni}
P_CI_Muni <-
	ggplot(data = sir_MGA) +
	geom_pointrange(aes(x= Muni, y = est, ymin = lower, ymax = upper, color = sig)) +
	labs(x = "Municipalities", y = "Standardized Morbidity Ratio") +
	theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=16)) +
	geom_hline(yintercept = 1, color = "Black") +

	scale_color_discrete(guide = guide_legend(title = "Significance")) +
	theme(legend.position="top") +
	theme(axis.text.x = element_text(size = 6,
																	 color = ifelse(sir_MGA$sig == "Significantly Higer",
																	 							 "green", "darkgrey")))
P_CI_Muni
```

# load again map to have clean one
## load Puerto Rico Municipality map
```{r loadMuniMap_Again, echo=FALSE, results='hide'}
load("/Spatial_Polygons/Puerto_Rico/PR_Muni_Map.RData")
```

```{r map2}
map2 <- merge(x = PR_Muni_Map, y = sir_MGA, by.x = "IDnum", by.y = "IDnum" )
head(map2)
```


```{r SigMap}
Sig_Map <- 
	ggplot(data = map2) +
	geom_polygon(aes(x = long, y = lat, group = id, fill = sig)) +
	
	geom_polygon(aes(x = long, y = lat, group = group, label = "id"),
								 data = map, 
							 colour = "black", fill = "grey", alpha = .1) + #
	theme(legend.position="bottom") +
	
		#geom_text(data = MuniSir2, aes(x = long, y = lat, label = round(sirMga, 1))) +
	
		# HospL
	geom_point(data = HospL,
						 aes(x = LONGITUDE, y = LATITUD),
						 color = "Yellow") +
# C330L
# 	geom_point(data = C330L,
# 						 aes(x = LONG, y = LAT),
# 						 color = "Blue") +
# CDTL	
# 	geom_point(data = CdtL,
# 						 aes(x = LONG_, y = LAT),
# 						 color = "Red") +
	
	coord_map() +
	
		scale_fill_brewer(palette = 5, guide = guide_legend(title = "Significance")) +
	
#	ggtitle("Alzheimer age-sex adjusted Standardized Morbidity Ratio by municipality, 2008-2013 \n Dots: Yellow = Hospitals") +
#; Blue = 330 Centers; Red = CDT"
	geom_text(data = NULL, x = -65.35, y = 18.46, 
						label = paste("N = ", format(nrow(AlDf),
																				 big.mark = ",")
						)) +
	labs(x = "Longitude", y = "Latitude") +
	geom_text(data = NULL, x = -65.50, y = 18.00, 
						label = paste("mean = ", round(mean(sir_MGA$est), 2),
													" ; ",
													"variance = ", round(var(sir_MGA$est), 2),
													sep = ""), size = 3) 
	
Sig_Map
```

```{r SaveObjects}
save(MapMuni_Crude_Rate, MapMuni_un_SIR, MapMuni_adjG_SIR,
		 MapMuni_adjGA_SIR, P_CI_Muni, Sig_Map,
file = "/Objects/Maps_SIR_Objects.Rdata")
```


