---
title: "Population Reshape for Analysis"
author: "José A. Bartolomei-Díaz, PhD - Outcome Project"
date: "July 11, 2015"
output: html_document
---

This document contains the procdures to obtain population distribution by municipality, gender and age group to be able to calculate the SIR.

# Generate Population Aggregation for the year 2010

## Load population by municipality, Gender and Age Group

```{r Load_popMuniSexAge}
Pop_Est <- read.csv2(file =
								 	"/media/truecrypt2/ORP2/Population_Estimates/Census/PR_Population_Estimates.csv",
					header = TRUE, sep = ",")

# Remove municipio Characters from the MUNICIPIO variable
Pop_Est <- data.frame(Pop_Est, Muni = 
		gsub(pattern = "Municipio", Pop_Est$NAME, replacement = "")
	)
```

### Reformat muni variable as character
```{r}
Pop_Est$Muni <- as.character(Pop_Est$Muni)
is.character(Pop_Est$Muni)
```

# Data set for manipulation
```{r newDF}
pop <- Pop_Est
```

# Subset data
Here we will subset the data and prepare reshaped data to be use in analyasis. 
The compleate work is to modify this script so all the years with estimates are reshaped not only the year 2010.

```{r ColSubset}
# Select a some columns of the original data set
pop <- pop[, c("Muni", "SEX", "AGEGRP", "POPESTIMATE2010")]

# Rename the column names
colnames(pop) <- c("Muni", "SEX", "AGEGRP", "EST10")

# Add labels to Sex variable
pop$SEX <- factor(pop$SEX, labels = c("Total", "Male", "Female"))

# Add laberls to Age Group Variables
pop$AGEGRP <- factor(pop$AGEGRP,
										 labels = c("Total", "0-4", "5-9", "10-14", "15-19",
										 					 "20-24", "25-29", "30-34", "35-39", "40-44",
										 					 "45-49", "50-54", "55-59", "60-64", "65-69",
										 					 "70-74", "75-79", "80-84", "85+"))

head(pop)
```

### Remove Rows labeled as Total
```{r}
pop <- pop[pop$SEX != "Total" & pop$AGEGRP != "Total",]

pop$SEX <- factor(pop$SEX)
pop$AGEGRP <- factor(pop$AGEGRP)
table(pop$SEX)
table(pop$AGEGRP)
head(pop)
```
# Create an 10 year age groups
```{r Age10}
pop$Age10 <- NA
# Age10 <- "0-19"
pop$Age10[pop$AGEGRP %in% c("0-4", "5-9", "10-14", "15-19")] <- "0-19"
 # Age10 <- "20-29"
pop$Age10[pop$AGEGRP %in% c("20-24", "25-29") ] <- "20-29"
# Age10 <- "30-39"
pop$Age10[pop$AGEGRP %in% c("30-34", "35-39") ] <- "30-39"
# Age10 <- "40-49"
pop$Age10[pop$AGEGRP %in% c("40-44", "45-49") ] <- "40-49"
# Age10 <- "50-59"
pop$Age10[pop$AGEGRP %in% c("50-54", "55-59") ] <- "50-59"
# Age10 <- "60-69"
pop$Age10[pop$AGEGRP %in% c("60-64", "65-69") ] <- "60-69"
# Age10 <- "70-79"
pop$Age10[pop$AGEGRP %in% c("70-74", "75-79") ] <- "70-79"
# Age10 <- "80+"
pop$Age10[pop$AGEGRP %in% c("80-84", "85+") ] <- "80+"

```
# Aggregate data by Age10 (I will over wirte the previous pop with this 10 year age aggregation)
```{r}
pop <- aggregate(EST10 ~ Muni + SEX + Age10, data = pop, FUN = sum)

# verify recode and aggregation
# aggregate(EST10 ~ AGEGRP, data = pop, FUN = sum)
# aggregate(EST10 ~ Age10, data = pop10, FUN = sum)
```


# The end result is an arry of popultion total of Municipality by Gender and Age group

```{r popMGA}
Pop_MGA <- pop
rm(pop)
head(Pop_MGA)
```

## Reshape population information as require functioin epi.inderectadj

```{r PopMG}
library(reshape2)
# 2010 Census Population by Municipality and Gender
Pop_MG <- dcast(Pop_MGA, formula = Muni ~ SEX, fun.aggregate = sum, value.var= "EST10")
head(Pop_MG)
```
# 2010 Census Population by Municipality and Age Group
```{r pMA}
Pop_MA <- dcast(Pop_MGA,
						 formula = Muni ~ Age10, fun.aggregate = sum,
						 value.var= "EST10")
head(Pop_MA)
```

# 2010 Census Population by Municipality, Gender and Age Group but casted
```{r pMGAc}
Pop_MGAc <- dcast(Pop_MGA,
						 formula = Muni ~ SEX + Age10, fun.aggregate = sum,
						 value.var= "EST10")
head(Pop_MGAc)

# Modify order of the columns to match observed cases data
Pop_MGAc <- data.frame(Muni = Pop_MGAc[, 1], Pop_MGAc[, c(10:17)], Pop_MGAc[, c(2:9)])
```

# Population by Gender
```{r Pop_Gen}
Pop_Gen <- aggregate(EST10 ~ SEX, data = Pop_MGA, FUN = sum)

head(Pop_Gen)
```

# Population by Age
```{r Pop_Age10}
Pop_Age10 <- aggregate(EST10 ~ Age10, data = Pop_MGA, FUN = sum)

head(Pop_Age10)
```

# Population by Municipality
```{r Pop_Muni}
Pop_Muni <- aggregate(EST10 ~ Muni, data = Pop_MGA, FUN = sum)

head(Pop_Muni)
```

# Save 2010 Estimates
```{r Save_2010_Estimates}
save.image("/media/truecrypt2/ORP2/Population_Estimates/Census/Pop_Estimates_2010.RData")
```

